---
title: "Lab_Assignment_5"
author: "Yifei Ding"
date: "2023-11-17"
output: html_document
---

```{r}
Sys.setenv(LANG = "en")
setwd("E:/R/R program/Data-Analysis-for-Economics-ECON106A-23-FALL/Lab Assignments")
library(tidyverse)
library(sf)
library(units)
```

Part I. Cleanup 1. Read in the District Areas Shapefile. Transform the CRS to CA Albers Equal Area.

```{r message=FALSE}
School_Districts<-st_read("DistrictAreas2122.shp")
```

2.  Read in the District College Outcomes csv file. Convert the CDCode to character. Create a variable that represents the CSU going rate. Create a variable that represents the UC going rate.
3.  Left Join the District Areas sf object to the District College outcomes.

```{r message=FALSE}
#read in college outcomes by district
District_College_Outcomes<-read_csv("District College Outcomes.csv")
District_College_Outcomes<-District_College_Outcomes%>%
  mutate(CDCode=as.character(CDCode))

School_Districts_join<-left_join(x=School_Districts, y=District_College_Outcomes)

#use albers projections
School_Districts_Albers<-st_transform(School_Districts_join, crs=3310)

School_Districts_rates<- School_Districts_Albers %>%
  mutate(CSU_rate=`Enrolled CSU (12 Months)`/`High School Completers`,
         UC_rate = `Enrolled UC (12 Months)`/`High School Completers`)
```

4.Read in the college locations csv file. Convert the data frame to a sf object. Transform the CRS to CA Albers Equal Area. Create a sf object for campuses that are part of the CSU system. Create a sf object for campuses that are part of the UC system

```{r message=FALSE}
#load excel file of GPS coordinates of college campuses into R
college_data <- read_csv("college locations.csv")

#select CSU's from this dataset
CSU_data <-college_data %>%
  filter(F1SYSNAM == "California State University")

# Convert CSU_data to a sf object
CSU_sf <- st_as_sf(CSU_data, coords = c("LONGITUD", "LATITUDE"), crs = 4269)

#select UC's from this dataset
UC_data <-college_data %>%
  filter(F1SYSNAM == "University of California")%>%
  filter(IALIAS!="UCOP")

# Convert CSU_data to a sf object
UC_sf <- st_as_sf(UC_data, coords = c("LONGITUD", "LATITUDE"), crs = 4269)

CSUSB<-CSU_sf%>%
  filter(IALIAS=="Cal State San Bernardino")
```

```{r}
#project to CA Albers

UC_Albers<-st_transform(UC_sf, crs=3310)

CSU_Albers<-st_transform(CSU_sf, crs=3310)

CSUSB_Albers<-st_transform(CSUSB, crs=3310)
```

Part II. Analysis 1. Create a map of CSU going rates by district with the CSU locations overlayed.

```{r}
#csu going & CSU locations
ggplot()+
  geom_sf(data=School_Districts_rates, aes(fill=`CSU_rate`))+
  geom_sf(data=CSU_Albers, color='red')
```

2.Create a map of the UC going rates by district with the UC locations overlayed.

```{r}
#uc going & uc locations
ggplot()+
  geom_sf(data=School_Districts_rates, aes(fill=`UC_rate`))+
  geom_sf(data=UC_Albers, color='red')
```

4.  Let's focus on school districts in Riverside County. Find the following:

```{=html}
<!-- -->
```
a.  The largest school district by area (sq miles)
b.  The largest school district by #enrolled students
c.  The school district with the highest CSU going rate

```{r}
Riverside_Districts<-School_Districts_rates%>%
  filter(CountyName=="Riverside")
```

```{r}
#what is the largest school district in Riverside by Area (square miles)?
Riverside_Districts%>%
  mutate(area=set_units(st_area(Riverside_Districts), mi^2))%>%
arrange(desc(area))%>%
  select(DistrictNa, area)%>%
  slice_head(n=5)
```

```{r}
#what is the largest school district in Riverside by # Total Enrollment?
Riverside_Districts%>%
  arrange(desc(EnrollTota))%>%
  select(DistrictNa, EnrollTota)%>%
  slice_head(n=5)
```

```{r}
#what is the school district with the highest CSU college going rates?
Riverside_Districts%>%
  arrange(desc(CSU_rate))%>%
  select(DistrictNa, CSU_rate)%>%
  slice_head(n=5)
```

5.  Calculate the distance between the Riverside School District Centroids and CSU San Bernadino (the closest CSU). Is the school district with the highest CSU going rate the closest to CSUSB?

```{r}
Jurupa_Unified<-Riverside_Districts%>%
  filter(DistrictNa=="Jurupa Unified")

set_units(st_distance(Jurupa_Unified, CSUSB_Albers), mi)
```

```{r}
Riverside_Districts_Centroids <- st_centroid(Riverside_Districts)

Distance_RDC_CSUSB <- set_units( st_distance(CSUSB_Albers, Riverside_Districts_Centroids) , mi)
colnames(Distance_RDC_CSUSB) <- Riverside_Districts_Centroids$DistrictNa
Distance_RDC_CSUSB
```

```{r}
data_q6 <- School_Districts_rates %>%
pivot_longer(cols= c(`CSU_rate`, `UC_rate`), names_to = "University_Systems", values_to = "College_Going_Rate")

data_q6 <- data_q6 %>% 
  group_by(CountyName,University_Systems) %>%
  summarise(College_going_rate_median = median(College_Going_Rate, na.rm = TRUE))
```

```{r}
data_q6 %>% na.omit() %>%
ggplot(mapping=aes(x= CountyName, y= College_going_rate_median)) +
geom_col(aes(color= University_Systems))+
scale_fill_brewer(palette = "Set2") +
theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.8))
```
